BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2015-12-22 Paper2/BOAST.tex
Added preliminary IEEE version.

\documentclass%[review]
              {elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}

@@

\journal{Journal of Parallel and Distributed Computing}

@@

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{frontmatter}

@@

%% or include affiliations in footnotes:
\author[mymainaddress]{Brice Videau}
\ead{brice.videau@imag.fr}
\author{Kevin Pouget}
\author{Luigi Genovese}
\author{Thierry Deutsch}
\author{Dimitri Komatitsch}
\author{Jean-François Méhaut}

\address[mymainaddress]{LIG/CNRS}

@@

\begin{keyword}
Code Generation \sep Portability \sep Genericity \sep Productivity and Software
Design \sep High Performance Computing \sep Autotuning \sep Non Regression
Testing
\end{keyword}

\end{frontmatter}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
\caption{Structure and Workflow of the BOAST framework.}
\label{fig:boast_workflow}
\end{center}
\end{figure}

@@

\subparagraph{Porting kernels to BOAST}

@@

\subparagraph{Porting run-time to \OCL} The second part of the port
consisted in the translation of the CPU-side of the application, from
\Cuda API to \OCL API. Most of the functions of the interfaces are
very similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of \OCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations, and 270 releases.

@@

\subparagraph{Debugging \OCL Execution: \code{GPUTrace}}

@@

\subparagraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
merged in \Specfem's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and N Intel
Xeon processors, we measured similar results between the original
\Cuda version and our BOAST-\Cuda version. With the same set of
optimization flags, BOAST \Cuda and \OCL version reported similar
execution time spans. The best execution speed was achieved with \Cuda
version though (25\% higher than \OCL), as one optimization parameter
(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
1.1. This parameter, in addition to specifying the work group size (which can
be done in \OCL), also constrains the number of work group that must run in
parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
the compiler must be very conservative on register usage in order to allow this
parallelism which allows better overlapping of communications end computations.
This functionality is not supported in \OCL.

>>>>

\documentclass{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}

@@

\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
%\journal{Journal of Parallel and Distributed Computing}

@@

%% `Elsevier LaTeX' style
\bibliographystyle{IEEEtran}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%\begin{frontmatter}

@@

%% or include affiliations in footnotes:
\author{Brice Videau, Kevin Pouget, Luigi Genovese, Thierry Deutsch, Dimitri Komatitsch, Frédéric Desprez, Jean-François Méhaut}

%\address[mymainaddress]{LIG/CNRS}

%\begin{keyword}
%Code Generation \sep Portability \sep Genericity \sep Productivity and Software
%Design \sep High Performance Computing \sep Autotuning \sep Non Regression
%Testing
%\end{keyword}

%\end{frontmatter}
\maketitle

@@

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{BOAST_Workflow.pdf}
\caption{Structure and Workflow of the BOAST framework.}
\label{fig:boast_workflow}
\end{center}
\end{figure}

@@

\paragraph{Porting kernels to BOAST}

@@

\paragraph{Porting run-time to \OCL} The second part of the port
consisted in the translation of the CPU-side of the application, from
\Cuda API to \OCL API. Most of the functions of the interfaces are
very similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of \OCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations, and 270 releases.

@@

\paragraph{Debugging \OCL Execution: \code{GPUTrace}}

@@

\paragraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
merged in \Specfem's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and N Intel
Xeon processors, we measured similar results between the original
\Cuda version and our BOAST-\Cuda version. With the same set of
optimization flags, BOAST \Cuda and \OCL version reported similar
execution time spans. The best execution speed was achieved with \Cuda
version though (25\% higher than \OCL), as one optimization parameter
(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
1.1. This parameter, in addition to specifying the work group size (which can
be done in \OCL), also constrains the number of work group that must run in
parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
the compiler must be very conservative on register usage in order to allow this
parallelism which allows better overlapping of communications end computations.
This functionality is not supported in \OCL.
